---
title: "notebook20161121"
author: "Mark Hagemann"
date: "November 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Start with current Pepsi and list of issues.

1. Many rivers had highly negative dA values. I get around this by subtracting the minimum from each dataset's dA.
2. Several rivers have zero-slopes for some cross-sections. I omitted these cross-sections from the computation.
3. Rivers with a lot of cross-sections (e.g. Mississippi) are very slow to run, and I just grabbed a subset of cross-sections when running the algorithm.
4. Many rivers are missing all but a single observed Q value. These are: MS (DS and US), Kanawha, Sacramento US, St. Lawrence (US and DS), Wabash
5. Kanawha widths are unrealistically large (>10km)
6. Platte is missing all dA values. Presumably because it is also missing 5 H values.
7. Tanana is missing Q for all but one cross-section, and the one cross-section (number 1) that does have Q extends the Q series beyond the times where it has W, S, etc. observations, resulting in missing values for these.
8. Connecticut doesn't work for the same reasons as before.

Tackle issues in this order: 4, 5, 6, 7

```{r}
badQrivs <- c("MississippiDownstream", "MississippiUpstream", "Kanawha",
              "SacramentoUpstream", "StLawrenceDownstream", 
              "StLawrenceUpstream", "Tanana", "Wabash")
PepsiNames <- sapply(Pepsi1, function(pc) pc[[1]][[1]])
badQinds <- match(badQrivs, PepsiNames)

i <- 0
```

```{r}
i <- i + 1

list.files(sprintf("data/%s", badQrivs[i]))
```

All except Sacramento upstream have .mat files.

```{r}
qfiles <- c("LowerMiss.mat", "UpperMiss.mat", "Kanawha.mat", 
            "upsacQ.xlsx", "downstreamData.mat", 
            "upstreamData.mat", "alltruth7.txt", "Wabash.mat")
qpaths <- paste("data", badQrivs, qfiles, sep = "/")
```

See how the tree-structured data looks

```{r}
Pepsi1[badQinds[1]]
```


```{r}
i <- i + 1
rivi <- badQinds[i]
# foo <- getPepsiData(Pepsi1[[rivi]])

Pepsi1[[rivi]]%>% 
  str()

# sapply(Pepsi0[[1]], length) # they're all length 1 -- nothing hiding.
```

So I need to add a 2-D numeric matrix with space down, time across.

```{r}
qpaths
qmats <- grepl("\\.mat$", qpaths)
origs <- lapply(qpaths[qmats], readMat)

getQ <- function(path, qname = "Qall") {
  if (!grepl("\\.mat$", path))
    return(NA)
  out0 <- readMat(path)
  out1 <- out0[[qname]]
  out1
}

qnames <- c(rep("Qall", 3), "Q", "Q", "Qall")
origQ <- mapply(getQ, path = qpaths[qmats], qname = qnames)

toRotate <- c(1:3, 6)
origQ[toRotate] <- lapply(origQ[toRotate], t)
lapply(origQ, class)

lapply(origQ, dim)

putQ <- function(Pepsi1_, riverind, qmatrix) {
  # qmatrix <- t(qmatrix)
  Pepsi1_[[riverind]][[1]][,,1]$Q <- qmatrix
  out <- Pepsi1_
}

foo <- putQ(Pepsi1_ = Pepsi1, riverind = 15, qmatrix = origQ[[1]])
```

Now replace each

```{r}
# i <- i + 1
# qpaths
# badQinds
Pepsi2 <- Pepsi1
for (i in 1:length(badQinds)) {
  Pepsi2 <- putQ(Pepsi1_ = Pepsi2, riverind = badQinds[i], qmatrix = origQ[[i]])
  print(i)
}

```


```{r}
plotQ <- function(Pepsi1_) {
  
  gdat <- lapply(Pepsi1_[badQinds], getPepsiData) %>% 
    setNames(badQrivs) %>% 
    bind_rows(.id = "river") %>% 
    arrange(time, xs, river) %>% 
    filter(xs == 1)
  
  out <- ggplot(gdat, aes(x = time, y = Q)) + 
    geom_line() + 
    facet_wrap(~river, scales = "free")
  out
}
```

```{r}
plotQ(Pepsi2)

```

Great! now for Sacramento

```{r}
sacq <- read.csv("data/SacramentoUpstream/upsacQ.csv") %>% 
  as.matrix() %>% 
  unname()
dim(sacq)

Pepsi2 <- putQ(Pepsi1_ = Pepsi2, riverind = badQinds[4], qmatrix = sacq)

Pepsi2[[badQinds[4]]][[1]][,,1] %>% str
```

plot again.

```{r}
plotQ(Pepsi2)
```

Awesome. Now put back into .mat form.

```{r}
Pepsi0 %>% str(2)

Pepsi2.5 <- list(All.pepsi.data = Pepsi2)
writeMat(con = "data/Pepsi_fixedup", data = Pepsi2.5)

foo <- readMat("data/Pepsi_fixedup")
str(foo, 1)
```

Not quite there...

```{r}
Pepsi3 <- setNames(Pepsi2, PepsiNames)
# writeMat("pepsi3test.mat", all.pepsi.data = Pepsi3)

# foo <- readMat("pepsi3test.mat")
# str(foo, 1)
```

Some googling turned up a package on github

```{r}
devtools::install_github("schmidtchristoph/reach/reach")
library(reach)
?matlabExportList
```

Try it!

```{r}
Pepsi3.mat <- matlabExportList(Pepsi2)
```

Doesn't work with nested lists. I may have to go a different route, like writing to and reading from json.

First tackle remaining items. 

### 5. Kanawha widths

```{r}
list.files("data/Kanawha")
kanawha <- readMat("data/Kanawha/Kanawha.mat")

str(kanawha, 1)
summary(unlist(kanawha$Wall))
class(kanawha$Wall)
summary(as.vector(kanawha$Wall))
summary(as.vector(kanawha$Qall))
```

Not too large here...

```{r}
Pepsi %>% head()

Pepsi %>% 
  filter(name == "Kanawha") %>% 
  summary()

Pepsi %>% 
  filter(name == "Kanawha") %>% 
  plot(W ~ time, data = .)
```

```{r}
badQrivs
badQinds
Pepsi3[[13]][[1]][,,1] %>% str

Pepsi4 <- Pepsi3
Pepsi4[[13]][[1]][,,1]$W <- t(kanawha$Wall)
Pepsi4[[13]][[1]][,,1] %>% str
```

Good. Now Platte dA

### 6. Platte problem areas


```{r}
list.files("data/Platte")
platte <- readMat("data/Platte/format_platte.mat")
plattru <- read.csv("data/Platte/truth.txt", sep = "\t")
head(plattru)

str(platte, 1)

Pepsi %>% 
  filter(name == "Platte") %>% 
  summary()


```

I should come up with a function to read truth.txt files.

But do this later. Easy fix first.

```{r}
?readLines
foo <- readLines(con <- file("data/Platte/truth.txt"))
close(con)
head(foo)
str(foo, 1)
foo

headLines <- grep("[[:alpha:]]", foo)
headLines
foo[headLines]

headLines[c(6, 7)] # everything between here is dA.

headLines[c(7, 11)] # everything between here is height.

foo[38:51]
```



```{r}
Pepsi %>% 
  filter(name == "Platte") %>% 
  select(time, xs, dA, H)
```

I should be able to do this all from Pepsi.

```{r}
calcdA <- function(width, height) {
  rnks <- rank(height, na.last = TRUE)
  ords <- order(height, na.last = TRUE)
  width <- width[ords]
  height <- height[ords]
  dH <- c(0, diff(height))
  delA <- width * dH
  
  out <- cumsum(delA)[rnks]
  out
}

cdatest <- Pepsi %>% 
  filter(name == "Po") %>% 
  group_by(xs) %>% 
  mutate(dA_myway = calcdA(W, H))
plot(dA_myway ~ dA, cdatest)

w1 <- cdatest$W[cdatest$xs == 1]
h1 <- cdatest$H[cdatest$xs == 1]

pladat <- Pepsi %>% 
  filter(name == "Platte", xs == 14) %>% 
  mutate(rand = rnorm(nrow(.))) %>% 
  arrange(rand)

pladat$H
with(pladat, calcdA(W, H))

```

That appears to work. So I'll add that in after the fact. 

### 7. Tanana Q

```{r}
list.files("data/Tanana")

platru <- readLines(con <- file("data/Tanana/alltruth7.txt"))
close(con)

headLines <- grep("[[:alpha:]]", platru)
headLines
platru[headLines]

headLines[c(5, 6)] # everything between here is Q.

platru[8:14]
plaq <- read.table(text = platru[8:14], sep = " ")
str(plaq)
plaqMat <- as.matrix(plaq)
str(plaqMat)
```

```{r}
badQrivs
badQinds
Pepsi4[[18]][[1]][,,1] %>% str
```

Alright, moving forward with readTruth function

```{r}

readTruth <- function
truthfile <- "data/Tanana/alltruth7.txt"
Aostart <- "A0"
Aoend <- "qt"
qstart <- "Qt"
qend <- "dA"
dAstart <- "dA"
dAend <- "h,"
hstart <- "h,"
hend <- "W,"
wstart <- "W,"
wend <- NA
tru <- readLines(con <- file(truthfile))
close(con)

readPiece <- function(startchars, endchars, trulines) {
  snum <- grep(paste0("^", startchars), trulines) + 1
  if (is.na(endchars)) {
    numlines <- grep("^[[:digit:]]", trulines)
    enum <- max(numlines)
  } else {
    enum <- grep(paste0("^", endchars), trulines) - 1
  }
  
  stopifnot(length(enum) == 1 && length(snum) == 1)
  
  piece <- read.table(text = trulines[snum:enum])
  out <- unname(as.matrix(piece))
  out
}

qpiece <- readPiece(startchars = qstart, endchars = qend, trulines = tru)
Aopiece <- readPiece(Aostart, Aoend, tru)
hpiece <- readPiece(hstart, hend, tru)
wpiece <- readPiece(wstart, wend, tru)

wt <- as.data.frame(t(wpiece))
ht <- as.data.frame(t(hpiece))
dApiece <- mapply(calcdA, width = wt, height = ht) %>% 
  t() %>% 
  as.matrix() %>% 
  unname()


putPiece <- function(Pepsi1_, riverind, what, newpiece) {
  if (!is(newpiece, "list"))
    newpiece <- list(newpiece)
  Pepsi1_[[riverind]][[1]][,,1][what] <- newpiece
  out <- Pepsi1_
}

Pepsi5 <- putPiece(Pepsi1_ = Pepsi4, riverind = 18, 
         what = c("W", "Q", "H", "Ao.calc", "dA"),
         newpiece = list(wpiece, qpiece, hpiece, Aopiece, dApiece))

str(Pepsi5[[18]][[1]][,,1])
```

Tanana has mismatched truth.txt, SWOTObs.txt dimensions.


That about wraps things up. Next just need to tidy up code and put into a script. 


### Plotting

```{r}
Pepsi_new %>% 
  filter(name != "Kanawha", name != "Tanana") %>% 
  group_by(name) %>% 
  filter(xs == sample(1:max(xs))) %>% 
  ungroup() %>% 
  ggplot(aes(x = time, y = Q)) + 
  geom_line() +
  facet_wrap(~name, scales = "free")

```



Hmm. NA shouldn't be a river.

```{r}
# i <- 1
Pepsi5[[i]][[1]][,,1] %>% str()
i <- i + 1
```


There are issues with:

- Kanawha (too many cross-sections)
- MississippiUpstream
- MississippiDownstream
- SacramentoUpstream
- Wabash
- Tanana




```{r}
str(Pepsi9$Kanawha, 1)
str(Pepsi6$Kanawha[[1]][,,1], 1)
str(Pepsi1[[13]][[1]][,,1], 1)

foo <- getPepsiData(Pepsi6$Kanawha)
summary(foo$xs)
unique(foo$xs)

foo <- getQ(path = "data/Kanawha/Kanawha.mat", qname = "Qall")
str(foo, 1)

foo <- readMat("data/Kanawha/Kanawha.mat")

list.files("data/StLawrenceDownstream")
stlmat <- readMat("data/StLawrenceDownstream/downstreamData.mat")
str(stlmat, 1)
stltru <- readTruth("data/StLawrenceDownstream/truth.txt")
str(stltru, 1)
str(Pepsi6$StLawrenceDownstream[[1]][,,1], 1)

stlpep <- getPepsiData(Pepsi6$StLawrenceDownstream)
summary(stlpep)

```

### Plotting, revisited

```{r}
Pepsi_new %>% 
  filter(name != "Kanawha", name != "Tanana") %>% 
  group_by(name) %>% 
  filter(xs == sample(1:max(xs))) %>% 
  ungroup() %>% 
  ggplot(aes(x = time, y = Q)) + 
  geom_line() +
  facet_wrap(~name, scales = "free")

```

```{r}
Pepsi_new %>% 
  filter(name != "Kanawha", name != "Tanana") %>% 
  group_by(name) %>% 
  filter(xs == sample(1:max(xs), 1)) %>% 
  ungroup() %>% 
  # select(H, dA, name, xs) %>% 
  ggplot(aes(x = H, y = dA)) + 
  geom_line() +
  facet_wrap(~name, scales = "free")

```


```{r}
Pepsi_new %>% 
  filter(name == "SacramentoUpstream") %>% 
  # group_by(name) %>% 
  filter(xs == sample(1:max(xs), 15)) %>% 
  # ungroup() %>% 
  # select(H, dA, name, xs) %>% 
  ggplot(aes(x = H, y = dA)) + 
  geom_line() +
  facet_wrap(~xs, scales = "free")

```

```{r}
Pepsi_new %>% 
  filter(name == "SacramentoUpstream") %>% 
  # group_by(name) %>% 
  filter(xs == sample(1:max(xs), 15)) %>% 
  # ungroup() %>% 
  # select(H, dA, name, xs) %>% 
  ggplot(aes(x = W, y = H)) + 
  geom_line() +
  facet_wrap(~xs, scales = "free")

```


Huh.

```{r}
p1 <- getPepsiData(Pepsi1[[grep("ntoDown", PepsiNames)]])
glimpse(p1)

p1 %>% 
  filter(xs == sample(1:max(xs))) %>% 
  # ungroup() %>% 
  ggplot(aes(x = H, y = dA)) + 
  geom_line() +
  facet_wrap(~name, scales = "free")

wfoo <- Pepsi1[[1]][[1]][,,1]$W[1,]
dafoo <- Pepsi1[[1]][[1]][,,1]$dA[1,]

plot(wfoo, dafoo)
```

I may need to fix up Sacramento dA.

```{r}
list.files("data/SacramentoUpstream")
list.files("data/SacramentoDownstream")

```



Looks like Q is being transferred wrong.

```{r}
foo <- getPepsiData(Pepsi6$Po)
plot(Q ~ time, foo)
```

That's good.

```{r}
Pepsi_new %>% 
  filter(name == "Po") %>% 
  plot(Q ~ time, .)
```

