---
title: "notebook20161121"
author: "Mark Hagemann"
date: "November 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Start with current Pepsi and list of issues.

1. Many rivers had highly negative dA values. I get around this by subtracting the minimum from each dataset's dA.
2. Several rivers have zero-slopes for some cross-sections. I omitted these cross-sections from the computation.
3. Rivers with a lot of cross-sections (e.g. Mississippi) are very slow to run, and I just grabbed a subset of cross-sections when running the algorithm.
4. Many rivers are missing all but a single observed Q value. These are: MS (DS and US), Kanawha, Sacramento US, St. Lawrence (US and DS), Wabash
5. Kanawha widths are unrealistically large (>10km)
6. Platte is missing all dA values. Presumably because it is also missing 5 H values.
7. Tanana is missing Q for all but one cross-section, and the one cross-section (number 1) that does have Q extends the Q series beyond the times where it has W, S, etc. observations, resulting in missing values for these.
8. Connecticut doesn't work for the same reasons as before.

Tackle issues in this order: 4, 5, 6, 7

```{r}
badQrivs <- c("MississippiDownstream", "MississippiUpstream", "Kanawha",
              "SacramentoUpstream", "StLawrenceDownstream", 
              "StLawrenceUpstream", "Wabash")
PepsiNames <- sapply(Pepsi1, function(pc) pc[[1]][[1]])
badQinds <- match(badQrivs, PepsiNames)
```

```{r}
i <- i + 1

list.files(sprintf("data/%s", badQrivs[i]))
```

All except Sacramento upstream have .mat files.

```{r}
qfiles <- c("LowerMiss.mat", "UpperMiss.mat", "Kanawha.mat", 
            "upsacQ.xlsx", "downstreamData.mat", 
            "upstreamData.mat", "Wabash.mat")
qpaths <- paste("data", badQrivs, qfiles, sep = "/")
```

See how the tree-structured data looks

```{r}
Pepsi1[badQinds[1]]
```


```{r}
i <- i + 1
rivi <- badQinds[i]
# foo <- getPepsiData(Pepsi1[[rivi]])

Pepsi1[[rivi]]%>% 
  str()

# sapply(Pepsi0[[1]], length) # they're all length 1 -- nothing hiding.
```

So I need to add a 2-D numeric matrix with space down, time across.

```{r}
qpaths
origs <- lapply(qpaths[-4], readMat)

getQ <- function(path) {
  if (!grepl("\\.mat$", path))
    return(NA)
  out0 <- readMat(path)
  out1 <- out0[["Qall"]]
  out1
}

origQ <- lapply(qpaths, getQ)

putQ <- function(Pepsi1_, riverind, qmatrix) {
  Pepsi1_[[riverind]][[1]][,,1]$Q <- qmatrix
  out <- Pepsi1_
}

foo <- putQ(Pepsi1_ = Pepsi1, riverind = 15, qmatrix = origQ[[1]])

```


```{r}
qfiles <- c("LowerMiss.mat", "UpperMiss.mat", "Kanawha.mat", 
            "upsacQ.xlsx", "downstreamData.mat", 
            "upstreamData.mat", "Wabash.mat")
qpaths <- paste("data", badQrivs, qfiles, sep = "/")


```

